Towards Comparable Active Learning

### 2023-06-01 doc/al_benchmark_paper.tex
paper update

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{(From Vijaya) Problem Description}
Given $n = l+u$ data points with $l \in \mathbb{N}$ many labeled examples $\mathcal{L} = \{(x_1, y_1),\ldots, (x_l,y_l)\}$, $u \in \mathbb{N}$ many unlabeled examples $\mathcal{U} = \{x_{l+1},\ldots,x_{n}\}$, a budget $\mathbb{N} \ni b \le u$ and an annotator $A: \mathbb{R}^M \to \mathbb{R}^C$ that can label $x$. We call $x \in \mathbb{R}^M$, $y \in \mathbb{R}^C$ predictors and labels respectively where $(x,y)$ are drawn from an unknown distribution $\rho$. 
Find an acquisition function $F: \mathcal{U},\mathcal{L}, b \mapsto \mathcal{U}_b$ that selects $b$ many examples from $\mathcal{U}$ such that $\mathcal{U}_b \subseteq \mathcal{U}$, $|\mathcal{U}_b|=b$ for labeling and the expected loss $\ell: \mathbb{R}^C \times \mathbb{R}^C \to \mathbb{R}$ (cross entropy loss) of the machine learning algorithm $\hat{y}$ (ex. classification) trained on $\mathcal{L} \cup \{\left(x, A(x)\right) \mid x \in \mathcal{U}_b\}$ is minimal: 
$$\min \quad \mathbb{E}_{(x,y) \sim \rho} \ell(y, \hat{y}(x))$$
Finding the optimal subset $\mathcal{U}_b$ is a combinatorial problem that is computationally not feasible for large $b$, so we allow sequential construction of the subset as a relaxation of the problem.
%To construct the active learning setting from a labeled dataset $\mathcal{D} = \{(x_1, y_1),\ldots, (x_n,y_n)\}$, we first split it into $\mathcal{D}_\text{train}$, $\mathcal{D}_\text{val}$ and $\mathcal{D}_\text{test}$ and then suppress the labels of $\mathcal{D}_\text{train}$ to form the unlabeled pool $\mathcal{U}$ and form and initial labeled pool by uniformly sampling $k$ number of instances per class, removing them from $\mathcal{U}$ and recovering their label $\mathcal{L} := \{(x_1, A(x_1)),\ldots, (x_{k*C},A(x_{k*C}))\} \quad s.t. \quad \mathcal{U} \cap \mathcal{L} = \emptyset$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{(From Lars) Problem Description}
Given two spaces $\mathcal{X}:=\mathcal{R}^M$ and $\mathcal{Y}:=\mathcal{R}^C$, a sample $\mathcal{D}_1,\ldots,\mathcal{D}_N \subseteq (\mathcal{X}\times \mathcal{Y})^*$ of sequences of pairs $(x,y)$  from an unknown distribution $p$ called datasets and a number $B\in\mathcal{N}$ with $B < |\mathcal{D}|$. \\
Given two functions $\ell:\mathcal{Y}\times \mathcal{Y}\rightarrow \mathcal{R}$ called loss, and $A: (\mathcal{X} \times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \mathcal{Y}^{\mathcal{X}}$ called learning algorithm, find a function
\begin{align*}
	a: (\mathcal{X}\times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \{0,1\}^*
\end{align*}
called acquisition function, s.t. the expected loss of a model learned on all predictors plus $B$ acquired targets is minimal.
%\begin{align*}
%	\mathbb{E}_{\mathcal{D}\text{train},\mathcal{D}\text{test}\sim p}   &
%	% \frac{1}{|\mathcal{D}\text{test}|} \sum_{(x,y)\in\mathcal{D}\text{test}}
%	% \operatorname{avg} \{
%	%  \ell(y, \hat y(x)) \mid (x,y)\in\mathcal{D}\text{test} \}
%	\operatorname{avg}\limits_{(x,y)\in\mathcal{D}\text{test}}
%	\ell(y, \hat y(x)) 
%	\\
%	\text{with }
%	\hat y:= & A( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_B}), \mathcal{D}\text{train}|_{\mathcal{X}})
%	\\ 
%	n_b := & \text{index}( a( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_{b-1}}), \mathcal{D}\text{train}|_{\mathcal{X}}) ),
%	\quad b\in 1{:}B
%\end{align*}
%(\textit{We would need to switch from lowest expected loss to highest AUC}) \\
Even though the acquisition function in principle could output the full subset of $\mathcal{D}_\text{train}$, the combinatorial problem is computationally not feasible for large $B$ and we allow sequential construction of the subset as a relaxation of the problem. \\
%
%To construct the active learning setting from existing labeled datasets, we first split it into $\mathcal{D}_\text{train}$, $\mathcal{D}_\text{val}$ and $\mathcal{D}_\text{test}$ and then suppress the labels $\mathcal{Y}$ of $\mathcal{D}_\text{train}$ to form the unlabeled pool $\mathcal{U} := \mathcal{X}^*$ and form and initial labeled pool by uniformly sampling $k$ number of instances per class from $\mathcal{U}$ and recovering their label $\mathcal{L} := (\mathcal{X}, \mathcal{Y})^{k*C}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}
We constrain our work on pool-based active learning where a pool of unlabeled samples is fixed at the start of each experiment and samples are chosen sequentially.
Specifically, we are not experimenting on so-called batch active learning, where at each iteration multiple unlabeled samples are chosen at the same time.
Even though batch AL is the more active research domain, it does not have a principled advantage over single-sample AL except speed of computation.
Not only is the problem of optimizing a portfolio of unlabeled samples more complicated to solve, the algorithms also have systematically less information per sample to work with.
For this reason we propose to focus more research effort on single-sample AL to find better algorithms in an environment that is simpler to solve and easier to control.
A performance comparison of batch AL and single-sample AL can be found in Fig. \ref{fig:ablation_bald}, which reproduces the message of Figure 1 from the paper \cite{gal2017deep} that proposed BALD, one of the SOTA algorithms for AL.
We can see that the batched version of BALD \cite{krizhevsky2009learning} can at most perform on-par with the single-sample algorithm.
Fig. \ref{fig:ablation_bald} also serves as a proof of concept for our provided code base.
Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature, as well as methodological AL papers with experiments on at least two data domains.
\begin{table}[h]
	\centering
	\begin{tabular}{l | c c c c c}
		Paper & Sampling & \# Datasets & Domains & Algorithms & Oracle \\
		\hline
		Beck et al. \cite{beck2021effective} & batch & 4 & 1 & 7 & - \\
		Hu et al. \cite{hu2021towards} & batch & 5 & 2 & 13 & - \\
		Li et al. \cite{li2022empirical} & batch & 5 & 1 & 13 & - \\
		Zhou et al. \cite{zhou2021towards} & batch & 3 & 2 & 2 & \checkmark \\
		\textbf{Ours} & single & 7 & 3 & 6 & \checkmark 
	\end{tabular}
	\caption{Comparison of our benchmark with the existing literature}
	\label{tab:benchmark_comparison}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
{\color{red} Version: Braindump}\\
Many different algorithms have been proposed for active learning. 
In this work we focus on those approaches that have shown consistent results over the years as well as some of the new approaches.
AL algorithms can be categorized into two classes: Geometric approaches and uncertainty-based approaches.
Geometric approaches use clustering techniques to partition the data and then sample their unlabeled points based on the clusters.
They often use the current classification model $\hat y_i$ to encode the data into a latent space to improve the performance of their clustering.
This benchmark includes the following geometric approaches: CoreSet \cite{sener2017active}, BADGE \cite{ashdeep} and TypiClust \cite{hacohen2022active}.
Uncertainty-based approaches use metrics to measure the classifiers state.
Commonly, a proxy for the sought after uncertainty of the model for a given datapoint is the distance of that point to the various decision boundaries, measured via the softmax output of the model.
This benchmark includes Shannon-Entropy sampling \cite{wang2014new}, margin sampling \cite{wang2014new} and BALD \cite{kirsch2019batchbald} \\ [1mm]
%
Some previous work also aimed to provide a benchmark suite for active learning:
The authors of \cite{beck2021effective} and \cite{li2022empirical} both focus on active learning in the image domain.
While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call ``Label Efficiency'' and provide experiments on many common configurations of data preparation, model training and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning to aid model training.
The authors of \cite{hu2021towards} study models that are learned with AL techniques in the image and text domain.
They test for several different properties of the models including robustness, response to compression techniques and final performance.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation}\label{sec:evaluation}
Following \cite{zhou2021towards}, the quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
We employ the normalized area under the accuracy curve (AUC):
\begin{equation}\label{eq:auc}
	\operatorname{AUC}(\mathcal{D}_{test}, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(\mathcal{D}_{test}, \hat y_i)
\end{equation}
where $\hat y_i$ is the (re-)trained classification model after the i-\textit{th} iteration.
To mimic the leave-one-out protocol for cross-validation we will restart each experiment multiple times.
Each restart will retain the train/test split (often given by the dataset itself), but introduces a new validation split.
The AUC incorporates performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
A good AL algorithm should be able to perform well in both scenarios. \\ [1mm]
Since AUC is dependent on the chosen budget, we need a general rule on how to set this hyperparameter that does not inherently benefit a subset of algorithms.
In this work, we choose the budget per dataset to be the first point at which any algorithm (except oracle) manages to reach a percentage of the upper bound performance measured on the full dataset.
Even though we would like to propose a single percentage value for all datasets, we found that different data modalities and use cases need different percentages to produce sensible budgets.
We propose the following values: \textbf{Tabular}: 99\%, \textbf{Image}: 90\% and \textbf{Text}: 95\%. \\ [1mm]
Additionally, we provide evidence in Fig. \ref{fig:restarts} that previous works have not evaluated their experiments with a sufficient number of restarts.
To create Fig. \ref{fig:restarts} we used 50 restarts from the margin/random sampling algorithm on the splice dataset.
From these 50 runs we uniformly sampled subsets of runs and calculated the median AUC for this subset.
One of these median AUC values corresponds to one cross-validated experiment sampled from the distribution of experiments that are restarted exactly this many times.
To create one slice in Fig. \ref{fig:restarts}, we drew 50 samples from this distribution.
Each box-plot represents the variance of an evaluation if conducted with the respective number of restarts.
We can observe that low repetitions ($<10$) provide an uncertain evaluation where lucky and unlucky draws of the same experiment give drastically different median AUC values.
\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{img/ablation_restarts.png}
	\caption{Random draws from an experimental distribution on the Splice datasets with different numbers of repetitions. Each point on the Y-axis represents a cross-validated result that could have been reported in a paper. This analysis shows the drastic differences in performance one could observe even when repeating an experiment 2-10 times.}
	\label{fig:restarts}
\end{figure}
To reliably arrive at the true median AUC, we propose to repeat every experiment 50 times, as only $>42$ repitions don't produce outliers anymore (as indicated by the rightmost columns in Fig \ref{fig:restarts}).
One way to reduce the number of necessary repetitions is to reduce the amount of variance in the experiment through specialized seeding (discussed in the next section).
We ultimately decided in favor of high variance and high number of repetitions as the high variance accurately reflects real world applications of AL.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Reproducibility}\label{sec:reproducibility}
A big focus in this work is to provide an experimental setup that is fully reproducible independent of the dataset, classification model or AL algorithm used.
For a fair comparison of two AL algorithms, both algorithms should receive equal starting conditions in terms of train / validation split, initialization of classifier and even the state of minor systems like the optimizer of mini-batch sampler.
Even though some implementations might have their own solution to some of these problems, to the best of our knowledge

%Decided against a fixed time budget
Given a seed, an evaluation on one dataset should always be done on the same validation split as well as the same random state for all included systems, like the mini batch sampler for model training or the initialization for the classifier itself.
Even though different AL algorithms will pick different samples, making them unavailable for sampling in earlier or later batches, the theoretical decision tree for every possible choice in every iteration $i$ should stay the same.
Since every possible trajectory cannot be precomputed and stored to disk, we need to resort to seeding.
The default choice of setting a global seed at the start of the experiment is not sufficient here, since a single additional random draw from the random number generator completely changes the behavior of all other systems.
This additional random number might be drawn during the initialization of the classification model or the AL algorithm, or even during every AL iteration if $\Omega$ is stochastic.
%This problem also applies to the initialization of the classification model $\theta$ and the initialization and querying of $\Omega$.
The desired control only be archived by assigning a separate random number generator to all these processes.
To the best of our knowledge, we are the first work that discusses this issue and proposes a solution for it.
We hypothesize that the insufficient setup with global seeds contributes to the on-going problem of inconsistent results of AL algorithms in different papers. \\ [1mm]
In summary, we introduce three different seeds: $s_\Omega$ for the acquisition function, $s_\mathcal{D}$ for dataset splitting and mini batch sampling and $s_\theta$ for model initialization and sampling of dropout masks.
Unless stated otherwise, we will keep $s_\Omega$ fixed for restarts of the same experiment, while $s_\mathcal{D}$ and $s_\theta$ are incremented by 1 between restarts to introduce stochasticity into our framework.
% This seeding setup enables a reproduction of the oracle curve even when the classifier or AL algorithm is different.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{(From Vijaya) Problem Description}
Given $n = l+u$ data points with $l \in \mathbb{N}$ many labeled examples $\mathcal{L} = \{(x_1, y_1),\ldots, (x_l,y_l)\}$, $u \in \mathbb{N}$ many unlabeled examples $\mathcal{U} = \{x_{l+1},\ldots,x_{n}\}$, a budget $\mathbb{N} \ni b \le u$ and an annotator $A: \mathbb{R}^M \to \mathbb{R}^C$ that can label $x$. We call $x \in \mathbb{R}^M$, $y \in \mathbb{R}^C$ predictors and labels respectively where $(x,y)$ are drawn from an unknown distribution $\rho$. 
Find an acquisition function $F: \mathcal{U},\mathcal{L}, b \mapsto \mathcal{U}_b$ that selects $b$ many examples from $\mathcal{U}$ such that $\mathcal{U}_b \subseteq \mathcal{U}$, $|\mathcal{U}_b|=b$ for labeling and the expected loss $\ell: \mathbb{R}^C \times \mathbb{R}^C \to \mathbb{R}$ (cross entropy loss) of the machine learning algorithm $\hat{y}$ (ex. classification) trained on $\mathcal{L} \cup \{\left(x, A(x)\right) \mid x \in \mathcal{U}_b\}$ is minimal: 
$$\min \quad \mathbb{E}_{(x,y) \sim \rho} \ell(y, \hat{y}(x))$$
Finding the optimal subset $\mathcal{U}_b$ is a combinatorial problem that is computationally not feasible for large $b$, so we allow sequential construction of the subset.
%To construct the active learning setting from a labeled dataset $\mathcal{D} = \{(x_1, y_1),\ldots, (x_n,y_n)\}$, we first split it into $\mathcal{D}_\text{train}$, $\mathcal{D}_\text{val}$ and $\mathcal{D}_\text{test}$ and then suppress the labels of $\mathcal{D}_\text{train}$ to form the unlabeled pool $\mathcal{U}$ and form and initial labeled pool by uniformly sampling $k$ number of instances per class, removing them from $\mathcal{U}$ and recovering their label $\mathcal{L} := \{(x_1, A(x_1)),\ldots, (x_{k*C},A(x_{k*C}))\} \quad s.t. \quad \mathcal{U} \cap \mathcal{L} = \emptyset$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{(From Lars) Problem Description}
Given two spaces $\mathcal{X}:=\mathcal{R}^M$ and $\mathcal{Y}:=\mathcal{R}^C$, a sample $\mathcal{D}_1,\ldots,\mathcal{D}_N \subseteq (\mathcal{X}\times \mathcal{Y})^*$ of sequences of pairs $(x,y)$  from an unknown distribution $p$ called datasets and a number $B\in\mathcal{N}$ with $B < |\mathcal{D}|$. \\
Given two functions $\ell:\mathcal{Y}\times \mathcal{Y}\rightarrow \mathcal{R}$ called loss, and $A: (\mathcal{X} \times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \mathcal{Y}^{\mathcal{X}}$ called learning algorithm, find a function
\begin{align*}
	a: (\mathcal{X}\times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \{0,1\}^*
\end{align*}
called acquisition function, s.t. the expected loss of a model learned on all predictors plus $B$ acquired targets is minimal.
%\begin{align*}
%	\mathbb{E}_{\mathcal{D}\text{train},\mathcal{D}\text{test}\sim p}   &
%	% \frac{1}{|\mathcal{D}\text{test}|} \sum_{(x,y)\in\mathcal{D}\text{test}}
%	% \operatorname{avg} \{
%	%  \ell(y, \hat y(x)) \mid (x,y)\in\mathcal{D}\text{test} \}
%	\operatorname{avg}\limits_{(x,y)\in\mathcal{D}\text{test}}
%	\ell(y, \hat y(x)) 
%	\\
%	\text{with }
%	\hat y:= & A( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_B}), \mathcal{D}\text{train}|_{\mathcal{X}})
%	\\ 
%	n_b := & \text{index}( a( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_{b-1}}), \mathcal{D}\text{train}|_{\mathcal{X}}) ),
%	\quad b\in 1{:}B
%\end{align*}
%(\textit{We would need to switch from lowest expected loss to highest AUC}) \\
Even though the acquisition function in principle could output the full subset of $\mathcal{D}_\text{train}$, the combinatorial problem is computationally not feasible for large $B$ and we allow sequential construction of the subset. \\
%
%To construct the active learning setting from existing labeled datasets, we first split it into $\mathcal{D}_\text{train}$, $\mathcal{D}_\text{val}$ and $\mathcal{D}_\text{test}$ and then suppress the labels $\mathcal{Y}$ of $\mathcal{D}_\text{train}$ to form the unlabeled pool $\mathcal{U} := \mathcal{X}^*$ and form and initial labeled pool by uniformly sampling $k$ number of instances per class from $\mathcal{U}$ and recovering their label $\mathcal{L} := (\mathcal{X}, \mathcal{Y})^{k*C}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}
We constrain our work on pool-based active learning where a pool of unlabeled samples is fixed at the start of each experiment and samples are chosen sequentially.
Specifically, we are not experimenting on so-called batch active learning, where at each iteration multiple unlabeled samples are chosen at the same time.
Even though batch AL is the more active research domain, it does not have a principled advantage over single-sample AL except speed of computation.
Not only is the problem of optimizing a portfolio of unlabeled samples more complicated to solve, the algorithms also have systematically less information per sample to work with.
For this reason we propose to focus more research effort on single-sample AL to find better algorithms in an environment that is simpler to solve and easier to control.
A performance comparison of batch AL and single-sample AL can be found in Fig. \ref{fig:ablation_bald}, which reproduces the message of Figure 1 from the paper \cite{gal2017deep} that proposed BALD, one of the SOTA algorithms for AL.
We can see that the batched version of BALD \cite{krizhevsky2009learning} can at most perform on-par with the single-sample algorithm.
Fig. \ref{fig:ablation_bald} also serves as a proof of concept for our provided code base.
Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature, as well as methodological AL papers with experiments on at least two data domains.
\begin{table}[h]
	\centering
	\begin{tabular}{l | c c c c c}
		Paper & Sampling & \#Datasets & \#Domains & \#Algorithms & Oracle \\
		\hline
		Beck et al. \cite{beck2021effective} & batch & 4 & 1 & 7 & - \\
		Hu et al. \cite{hu2021towards} & batch & 5 & 2 & 13 & - \\
		Li et al. \cite{li2022empirical} & batch & 5 & 1 & 13 & - \\
		Zhou et al. \cite{zhou2021towards} & batch & 3 & 2 & 2 & \checkmark \\
		\textbf{Ours} & single & 7 & 3 & 6 & \checkmark 
	\end{tabular}
	\caption{Comparison of our benchmark with the existing literature. Oracle curves serve as an approximation of the best possible AL algorithm.}
	\label{tab:benchmark_comparison}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
{\color{red} Version: Braindump}\\
\begin{wrapfigure}{r}{0.5\textwidth}
	\centering
	\includegraphics[width=\linewidth]{img/ablation_bald}
	\caption{Shows a reproduction of the performance analysis of BALD from the original authors \cite{gal2017deep} in comparison to entropy sampling and an adaptation of BALD for batch AL \cite{kirsch2019batchbald} with a batch size of 40.}
	\label{fig:ablation_bald}
\end{wrapfigure}
Many different algorithms have been proposed for active learning. 
In this work we focus on those approaches that have shown consistent results over the years as well as some of the new approaches.
AL algorithms can be categorized into two classes: Geometric approaches and uncertainty-based approaches.
Geometric approaches use clustering techniques to partition the data and then sample their unlabeled points based on the clusters.
They often use the current classification model $\hat y_i$ to encode the data into a latent space to improve the performance of their clustering.
This benchmark includes the following geometric approaches: CoreSet \cite{sener2017active}, BADGE \cite{ashdeep} and TypiClust \cite{hacohen2022active}.
Uncertainty-based approaches use metrics to measure the classifiers state.
Commonly, a proxy for the sought after uncertainty of the model for a given datapoint is the distance of that point to the various decision boundaries, measured via the softmax output of the model.
This benchmark includes Shannon-Entropy sampling \cite{wang2014new}, margin sampling \cite{wang2014new} and BALD \cite{kirsch2019batchbald} \\ [1mm]
%
Some previous work also aimed to provide a benchmark suite for active learning:
The authors of \cite{beck2021effective} and \cite{li2022empirical} both focus on active learning in the image domain.
While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call ``Label Efficiency'' and provide experiments on many common configurations of data preparation, model training and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning to aid model training.
The authors of \cite{hu2021towards} study models that are learned with AL techniques in the image and text domain.
They test for several different properties of the models including robustness, response to compression techniques and final performance.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation}\label{sec:evaluation}
Following \cite{zhou2021towards}, the quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
We employ the normalized area under the accuracy curve (AUC):
\begin{equation}\label{eq:auc}
	\operatorname{AUC}(\mathcal{D}_{test}, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(\mathcal{D}_{test}, \hat y_i)
\end{equation}
where $\hat y_i$ is the (re-)trained classification model after the i-\textit{th} iteration.
To mimic the leave-one-out protocol for cross-validation we will restart each experiment multiple times.
Each restart will retain the train/test split (often given by the dataset itself), but introduces a new validation split.
The AUC incorporates performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
A good AL algorithm should be able to perform well in both scenarios. 
Since AL has high variance and is prone to outliers, we propose to report median AUC instead of mean AUC values. \\ [1mm]
The AUC is dependent on the chosen budget, so we need a general rule on how to set this hyperparameter that does not inherently benefit a subset of algorithms.
In this work, we choose the budget per dataset to be the first point at which any algorithm (except oracle) manages to reach a percentage of the upper bound performance measured on the full dataset.
Even though we would like to propose a single percentage value for all datasets, we found that different data modalities and use cases need different percentages to produce sensible budgets.
We propose the following values: \textbf{Tabular}: 99\%, \textbf{Image}: 90\% and \textbf{Text}: 95\%. \\ [1mm]
Additionally, we provide evidence in Fig. \ref{fig:restarts} that previous works have not evaluated their experiments with a sufficient number of restarts.
To create Fig. \ref{fig:restarts} we used 50 restarts from the margin/random sampling algorithm on the splice dataset.
From these 50 runs we uniformly sampled subsets of runs and calculated the median AUC for this subset.
One of these median AUC values corresponds to one cross-validated experiment sampled from the distribution of experiments that are restarted exactly this many times.
To create one slice in Fig. \ref{fig:restarts}, we drew 50 samples from this distribution.
Each box-plot represents the variance of an evaluation if conducted with the respective number of restarts.
We can observe that low repetitions ($<10$) provide an uncertain evaluation where lucky and unlucky draws of the same experiment give drastically different median AUC values.
\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{img/ablation_restarts.png}
	\caption{Random draws from an experimental distribution on the Splice datasets with different numbers of repetitions. Each point on the Y-axis represents a cross-validated result that could have been reported in a paper. This analysis shows the drastic differences in performance one could observe even when repeating an experiment 2-10 times.}
	\label{fig:restarts}
\end{figure}
To reliably arrive at the true median AUC, we propose to repeat every experiment 50 times, as only $>42$ repitions don't produce outliers anymore (as indicated by the rightmost columns in Fig \ref{fig:restarts}).
One way to reduce the number of necessary repetitions is to reduce the amount of variance in the experiment through specialized seeding (discussed in the next section).
We ultimately decided in favor of high variance and high number of repetitions as the high variance accurately reflects real world applications of AL.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Reproducibility}\label{sec:reproducibility}
A big focus in this work is to provide an experimental setup that is fully reproducible independent of the dataset, classification model or AL algorithm used.
For a fair comparison of two AL algorithms, both algorithms should receive equal starting conditions in terms of train / validation split, initialization of classifier and even the state of minor systems like the optimizer or mini-batch sampler.
Even though some implementations might have their own solution to some of these problems, to the best of our knowledge no previous work has discussed this topic in detail.
At the core of this problem is the seeding utility in PyTorch, Tensorflow and other framework, whose default choice is a single global seed.
Since many system draw random numbers from this seed all of them influence each other to a point where a single additional draw can completely change the model initialization or data split, depending on the order in the implementation.
Even though you could find workarounds like re-setting the seed multiple times, this problem also extends to every AL iteration and the systems within.
We propose an implementation that creates a separate Random Number Generator (RNG) for each of these systems to ensure equal testing conditions even when the AL algorithm, dataset or classifier changes.
We hypothesize that the insufficient setup with global seeds contributes to the on-going problem of inconsistent results of AL algorithms in different papers. \\ [1mm]
In summary, we introduce three different seeds: $s_\Omega$ for the AL algorithm, $s_\mathcal{D}$ for dataset splitting and mini batch sampling and $s_\theta$ for model initialization and sampling of dropout masks.
Unless stated otherwise, we will keep $s_\Omega$ fixed for restarts of the same experiment, while $s_\mathcal{D}$ and $s_\theta$ are incremented by 1 between restarts to introduce stochasticity into our framework.
Some algorithms require a subsample to be drawn from $\mathcal{U}$ in order to reduce the computational cost in each iteration, while others need access to the full unlabeled pool (i.e. for effective clustering).
If a subsample is required, it will be drawn from $s_\Omega$ and therefore will not influence other systems in the experiments.
For each algorithm, we decided if subsampling is required based on our available hardware, but decided against setting a fixed time limit per experiment, since this would introduce unnecessary complexity into the benchmark.
An overview of selected hyperparameters per AL algorithm can be found in Appendix \ref{app:agent_hyperparameters}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Oracle}\label{sec:oracle}
Posing active learning as a combinatorial problem, the oracle set $\mathcal{U}_b$ for a given dataset, model and training procedure would be the set that induces the highest AUC score for a given budget.
However, since this problem is not solvable for realistic datasets, previous works have proposed approximations to this oracle sequence.
\cite{zhou2021towards} has used simulated annealing to search for the optimal sequence and used the best  solution found after a fixed time budget. 
Even though their reported performance curves display a significant lift over all other algorithms, we found the computational cost of reproducing this oracle for all our datasets to be prohibitive (The authors reported the search to take several days per dataset on 8 V100 GPUs).
\begin{algorithm}
	\caption{Oracle}\label{alg:oracle}
	\begin{algorithmic}[1]
		\Require $\mathcal{U}, \mathcal{L}, \mathcal{Y}, \mathcal{D}_\text{test}$ Train, Margin, $\tau, \hat y_\theta$ 
		\State $\text{acc} \gets \operatorname{Train}(\mathcal{L}, \mathcal{D}_\text{test}, \hat y_\theta)$ 
		%\State $u \underset{\tau}{\sim} \text{unif}(1:|\mathcal{U}|)$
		\State $r^* \gets 0$
		\For{$t := 1 \ldots \tau$}
		\State $\mathcal{L}' \gets \mathcal{L}^{(i)} \cup \{(\mathcal{U}_t, y_t)\}$
		\State $\text{acc}' \gets \operatorname{Train}(\mathcal{L}', \mathcal{D}_\text{test}, \hat y_\theta)$  
		\State $r \gets \text{acc} - \text{acc}'$
		\If{$r > r^*$} 
		\State $r^* \gets r$
		\State $u^* \gets \mathcal{U}_t$
		\EndIf
		\EndFor
		\If{$r^* = 0$}
		\State $u^* \gets \operatorname{margin}(\mathcal{U}, \hat y_\theta)$
		\EndIf
		\Return $u^*$
	\end{algorithmic}
\end{algorithm}
In this paper we propose a greedy oracle algorithm that constructs an approximation of the optimal sequence in an iterative fashion.
Our oracle simply evaluated every data point in the provided sample of unlabeled points by fitting the classifier and directly measuring the resulting test performance.
The point with the best test performance is selected and added to the labeled pool for that iteration.
We noticed that this oracle is overfitting on the test set, resulting in stagnating or even decreasing performance curves in later AL iterations.
To circumvent this problem, we introduced margin sampling as a fallback option for the oracle.
Whenever the oracle does not find an unlabeled point that results in an increase in performance (indicating an overfitting position), it defaults to margin sampling in that iteration.
The pseudocode for our oracle can be found in Alg. \ref{alg:oracle}.
In the algorithm $\operatorname{Retrain}(\mathcal{L}^{(i)}, \hat y_\theta)$ trains the classification model $\hat y_\theta$ and returns the accuracy on the test set $\mathcal{D}_\text{test}$. \\ 
$y_t$ is shorthand for the corresponding label of $u^{(i)}_{t}$ that can be recovered from the dataset labels. \\
When the oracle does not find a sample with positive change in classification performance ($r^* = 0$), it applies margin sampling as a fallback ($\operatorname{margin}(u^{(i)}, \hat y_\theta)$). \\
Alg. \ref{alg:oracle} replaces the algorithm in the AL process.
